---
title: "GWAS - outline"
author: "Filippo Biscarini"
date: "October 10, 2018"
output: html_document
bibliography: gwas.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("knitr")
library("dplyr")
library("knitr")
library("GenABEL")
library("ggplot2")
library("reshape2")
```

## R Markdown

GWAS for binary traits.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this @meuwissen2001prediction:

```{r data, echo=FALSE}
setwd("~/Dropbox/cursos/berlin2018")
load("data/df_genabel.RData")

descriptives.marker(df)
```

```{r, echo=FALSE}
kable(descriptives.trait(df))
```

N. `r gtdata(df)@nsnps` from `r length(unique(gtdata(df)@chromosome))` chromosomes

```{r, echo=FALSE}
dd <- as.data.frame(table(gtdata(df)@chromosome))
names(dd) <- c("chrom","nSnps")
kable(dd)
```

## Including Plots

For the moment, our "phenotype" coincides with the breed:

```{r pressure, echo=FALSE}
pp <- phdata(df)
pp$phenotype <- as.factor(pp$phenotype)
p <- ggplot(pp,aes(x=phenotype)) + geom_bar(aes(fill=phenotype))
p <- p + facet_wrap(~breed) + xlab("phenotype")
print(p)
```

### Quality check

```{r, label='quality_check', echo=FALSE}
qc1 <- check.marker(df, p.level=0)
df1 <- df[,qc1$snpok]
## phenotypes: exclude unknown status
df2 <- df1[phdata(df1)$phenotype!=0,]
```

### Kinship matrix

```{r, label='kinship', echo=FALSE}
K <- ibs(df2,weight = "freq")
K[upper.tri(K)] = t(K)[upper.tri(K)]
heatmap(K,col=rev(heat.colors(75)))
```

```{r, label='polygenic'}
ff <- phdata(df2)$phenotype
ff[ff==2] <- 0
df2 <- add.phdata(df2,ff, name = "pheno_recoded")

# h2a <- polygenic(pheno_recoded,data=df2,kin=K,trait.type = "binomial",fixh2 = 0.5,llfun = "polylik")
# df.mm <- mmscore(h2a,df1)
# descriptives.scan(df.mm,top=100)
# plot(df.mm,col = c("red", "slateblue"),pch = 19, cex = .5, main="phenotype")
```

```{r, label='eigenn_decomposition', echo=FALSE}
df.mm  <- egscore(phdata(df2)$phenotype,data=df2,kin=K)
plot(df.mm,col = c("red", "slateblue"),pch = 19, cex = .5, main="trait")
descriptives.scan(df.mm,top=10)
```


## References
